---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data
```{r setoptions, echo=TRUE}
data_file <- "activity.zip"
unzip(data_file)
data <- read.csv("activity.csv", header = TRUE, sep = ",")
```


## What is mean total number of steps taken per day?
```{r Mean}
TotalStepsPerDay <- tapply(data$steps, data$date, FUN = sum, na.rm = TRUE)
```

```{r histogram}
hist(TotalStepsPerDay, col = "purple", xlab = "Steps per day", breaks =10, main = "Total Steps Per Day")
```

```{r Mean and Median}
mean(TotalStepsPerDay, na.rm=TRUE)

median(TotalStepsPerDay, na.rm = TRUE)

```
## What is the average_daily_activity_pattern?
```{r Average_Daily_Activity}
averagePattern <- aggregate(x=list(steps=data$steps), by= list(Interval = data$interval), FUN = "mean" , na.rm=TRUE) 
plot(averagePattern$Interval,averagePattern$steps, type = "l", xlab= "5 minute Interval", ylab="Steps", main="Average Daily Activity Pattern")
```
# Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r max number of steps}
averagePattern[which.max(averagePattern$steps),]
```

## Imputing missing values

# Total number of missing values in the dataset 
```{r number of missing Values}
sum(is.na(data$steps))
```

# There are many missing values in data noted as NA. So, to fill the missing values we take the mean of that column and substitue NA with mean
```{r Missing Values}
data$steps[is.na(data$steps)] <- mean(data$steps, na.rm = TRUE)
```
# Create a histogram using new data with filled values
```{r histogram_after_Imputing_missing_values}
Steps_impute <- tapply(data$steps, data$date, FUN = sum)
hist(Steps_impute, xlab="Steps", main="Total Steps per day", breaks = 10, col = "orange")
```

```{r Mean and Median after imputing}
mean(Steps_impute)

median(Steps_impute)

```
Mean and Median Values increased after imputing missing values 

## Are there differences in activity patterns between weekdays and weekends?

```{r}
week <- function(date) {
        day <- weekdays(date)
       ifelse(day %in% c("Saturday", "Sunday"), "weekend", "weekday")
}
data$date <- as.Date(data$date)
data$day <- sapply(data$date, FUN = week)
```

Plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)
```{r time_series_plot}
average_day <- aggregate(steps~ interval+day, data = data , mean)
library(ggplot2)
ggplot(average_day, aes(interval,steps)) + geom_line() + facet_grid(day~.) + xlab("5-minute interval") + ylab("Number of steps")
```